Joint training of classification model and similarity model for low-resource text classification in chatbot

### 2020-09-08 chatbot_classification/aaai_2021/paper.tex
version 05.   modified:   paper.tex   new file:   paper.ver05.pdf

    \item   \emph{Non-pretrained   model  based  classification  system},  which
    consists  of  a popular CNN based TextCNN \cite{kim2014convolutional}, and a
    recently  proposed label-embedding based LEAM (todo: reference). Empircally,
    a  non-pretrained  model  based  system  are  not  performing  as  well as a
    pretrained  models  based,  if  both is applicable in some task. Our listing
    these  non-pretrained model based systems here is to provide a comprehensive
    performance comparison on our datasets.

@@

  \textbf{Comparison  among  SFC  variants.} First, we turn to compare among the
  four  SFC  variants shown in Table \ref{tbe:table2} to analyze the improvement
  brought  by  multi-task  and  joint-model  structure. We conduct
  ablation  experiments  to study the benefits brought by multi-task training by
  comparing multi-task two-stage SFC with single-task two-stage SFC. We can see from
  Table  \ref{tbe:table2}  that  two-stage  SFC  with  only  task 1 (sentence-pair
  similarity task) degrades by 0.80 percent on average and two-stage SFC with only
  task  2  (top-K  classification  task) degrades by 0.45 percent on average. The
  degradation  indicates  that  training  with these 2 tasks together in SFC can
  effectively  help  the  sentence-pair  model  extract  more knowledge from the
  limited  amount  of data. Besides, task 2 degrades 0.35 percent less than task
  1.  This observation suggests that task 2 may play a relatively more important
  role  in multi-task learning, which leads us to put larger weight $\alpha_{K}$
  to  the  loss  $\mathcal{L}^{K}$  of  task  2  when calculating the total loss
  $\mathcal{L}$.  Also, we can see that the joint SFC steadily outperforms
  two-stage  SFC  on  5  diverse  datasets  by  0.95  percents  on  average.  This
  observation  supports  the idea that the joint model structure can
  alleviate the limitation brought by the lack of interaction between similarity
  model  and  classification  model in two-stage SFC. Therefore, for the following
  analysis,  we  will  focus  on  the comparison between joint SFC and the
  other baseline models.

  \textbf{Analysis  under  few-shot  scenarios.}  Next,  we  analyze the overall
  improvement trend of SFC under various few-shot settings. Our main focus is to
  study chatbot building in common situation during real-world application where
  only  a  few  sample  sentences are available for each class. The experimental
  results  for  the  3  diverse  intention detection datasets (ClINC150, BANKING77,
  HWU64) under few-shot settings(5-20 samples per class) indicate that our proposed joint SFC can
  achieve  2.47  percents  average  improvement  over  one  of the most powerful
  baseline  models,  which  is  RoBERTa-large classification model. Moreover, the improvement in F-score becomes more and more
  prominent  with the decrease of data sample number per class. Our joint
  SFC   can  achieve  4.97  percents  average  improvements  over  RoBERTa-large
  classification model in the most extreme few-shot setting with only 5 training
  samples  per class. This observation indicates that SFC is especially well-performed
  when  being  applied during the initial stage of building task-specific chatbot
  system,  where  the  amount  of  data sample for each class label is extremely scarce. Furthermore, our proposed joint SFC can also achieve 2.04 percents average improvement over RoBERTa-large classification model on all the data settings including the ones with large scale (30-50 samples per class). This experimental result indicates that joint SFC can still steadily outperforms single-task pretrained model baseline even if the data size grows bigger. But the improvement is still more prominent when the data is scarce, which makes SFC an excellent model for low-resource chatbot scenario.

  \textbf{Comparison  between  label  embedding  and  sentence  pair embedding.}
  Afterwards,  we  start  to  compare the performance of SFC with the text-label
  joint  embedding  baseline  named as LEAM. As shown in Table \ref{tbe:table2},
  joint  SFC outperforms LEAM by 11.90 percents on average F-score over 5
  datasets.  Besides,  although the labels within the 3 diverse intention detection
  datasets  (CLINC150,  BANKING77,  HWU64)  are  all  being  well-defined as the
  semantic  summary  of  each intention, joint SFC still outperforms LEAM by
  10.71  percents  on  average  F-score.  This  observation  indicates that it is
  usually  quite  hard  to  use  a  fixed  label  with  only  a few words as the
  representative  of  all  the  sentences  within a class since there are always
  plenty of ways for people to express questions or queries about the same topic
  during   a   conversation  with  chatbot.  Therefore,  learning  the  semantic
  similarity  between  user  query  and  sample  sentences through sentence pair
  embedding  can be more reliable than making decision according to the semantic
  relationship between user query and class labels through label embedding.

  \textbf{Fusion  of  classification  and  similarity  model.} Then, we begin to
  compare  the  general  performance  of SFC with single-task classification and
  similarity  baselines.  As  for  the  classification  model,  joint SFC
  outperforms    TextCNN    by   21.52   percents,   ALBERT-xxlarge   by   7.99
  percents\footnote{We found that the performance of ALBERT.xxlarge is extremely
  unstable  for  classification  task.  Here, we did multiple run with different
  settings  of parameters such as learning rate and batch size for each dataset,
  and  record  the  best  result  we  can  obtain  in  Table \ref{tbe:table2}.},
  BERT-large  by  7.73 percents, RoBERTa-base by 3.79 percents, RoBERTa-large by
  2.04 percents on average F-score over 5 datasets. In this way, we can see that
  RoBERTa  is  the  best  baseline  pre-trained  model for the low-resource text
  classification  task  we  are  studying.  As for the similarity model, we only
  choose  the  strongest  pretrained  model  RoBERTa-large  as our baseline, and
  joint  SFC  achieves  the  improvement  of  7.09  percents  on average.
  Moreover,  joint  SFC's  result on 2 datasets without data sampling for
  few-shot  settings,  which  are  ITG  and  Amazon-670k, are also 1.24 and 2.89
  percents  higher  than  Roberta-large  classification  model, and are 6.96 and
  10.77  percents higher than Roberta-large similarity model. These improvements
  illustrate that putting classification model and similarity model at different
  network  layers  to  cooperate  with  each  other  is  a  successful design of
  joint network structure for joint model, which enables these two models
  assist each other for better overall results.

  \textbf{Settings  of  hyperparameters.} Finally, we analyze the effect brought
  by  different  settings  of  hyperparameters,  see  Table \ref{tbe:table3}. We
  conduct  experiments on 5 different settings of hyperparameters, which are the
  number  of candidate class number $K$ and the number of sentence pairs $P$ for
  each  class, on ITG and Amazon-670K datasets to find the most suitable setting
  of  these  2  hyperparameters.  We  control the total number of sentence pairs
  sampled  for  each  user  query  into  about  the same level of amount(50-60).
  Through  analyzing the experimenal results shown in Table \ref{tbe:table3}, we
  find  that the best setting of hyperparameters might be different for
  datasets of different data size. The class label number of ITG and Amazon-670k
  are  similar  in  our  experimental  setting, which is around 250. However the
  total  amount  of  data  points  of  ITG  is approximately 2 times larger than
  Amazon-670k. Therefore, the setting of $K=5$ and $P=10$ is good enough for ITG
  since  most  of  the  classes  contains  over  10  data  points.  However, for
  Amazon-670k,  the amount of data sample for each class is scarce, thus, we may
  need  more  candidate  classes  to  sample enough effective sentence pairs for
  sentence-pair model layer of SFC. In this way, the setting of $K=10$ and $P=5$
  becomes a good choice. However, in spite of the effect on evaluation performance brought by different settings of hyperparameters, joint SFC still steadily outperforms two-stage SFC by 0.41 percents in overall average F-score on ITG and Amazon-670k datasets.

>>>>

    \item   \emph{Non-pretrained   model  based  classification  system},  which
    consists  of  a popular CNN based TextCNN \cite{kim2014convolutional}, and a
    recently  proposed label-embedding based LEAM (todo: reference and its label
    setting). Empircally,
    a  non-pretrained  model  based  system  are  not  performing  as  well as a
    pretrained  models  based,  if  both is applicable in some task. Our listing
    these  non-pretrained model based systems here is to provide a comprehensive
    performance comparison on our datasets.

@@

  \subsubsection*{Which SFC variant performs best?} 
  First, we compare the four  SFC  variants to analyze the improvement
  brought  by  multi-task  and  joint-model  structure.

  Comparing with training two-stage SFC with multi-task, training with only task
  1,  corresponding  to  sentence pair similarity model, degrades by 0.8 percent
  point  on average, and training with only task 2, corresponding to top-K based
  classification  task,  degrades  by  0.45  percent  point  on  average.  These
  degradation  indicates  multi-task training in two-stage SFC is helpful for the
  system  performance.  Besides, task 2 play a relatively more important role in
  multi-task learning, and this aligns with their optimal weight settings.

  Joint SFC consistently outperforms two-stage SFC on 5 diverse datasets by 0.95
  percents  on  average. This observation supports the idea that the joint model
  structure  can  alleviate  the  limitation  brought by the lack of interaction
  between  two  stages.  In  following analysis, we will focus on the comparison
  between joint SFC and the other baseline models.

  \subsubsection*{Is  the  fusion of a classification model and similarity model
  helpful?}  We  compare  the performance of joint SFC with classification based
  and similarity based baselines.

  Regarding the  classification  models,  joint SFC
  outperforms    TextCNN    by   21.52   percent points,   ALBERT-xxlarge   by   7.99
  percent points, BERT-large  by  7.73 percent points, RoBERTa-base by 3.79 
  percent points, RoBERTa-large by 2.04 percents on average F-score over 5 datasets.  
  Especially, ALBERT.xxlarge based is much more  unstable  in our senario. Thus, 
  we did multiple run with different settings, e.g.,  learning rate and batch size 
  for each dataset, and  report the  best  result  here. Comparatively,  
  RoBERTa  based is  the  most stable and has the best performance among
  all baselines and all experiment settings.

  Regarding  the  similarity  model,  as  analyzed  above,  we  only  choose the
  RoBERTa-large based as our baseline. Our joint SFC achieves the improvement of
  7.09  percents  on average. Moreover, joint SFC's result on 2 datasets without
  data  sampling  for few-shot settings, which are ITG and Amazon-670k, are also
  1.24 and 2.89 percents higher than Roberta-large classification model, and are
  6.96 and 10.77 percents higher than Roberta-large similarity model (todo:
  meaning?).

  The above analysises illustrate that suitable fusing the classification model and the 
  similarity model to make them cooperate with each other is helpful to better
  performance.

  \subsubsection*{How does joint SFC interact with the classfication in stage
  1?}
  {todo: content}

  \subsubsection*{How does SFC perform with different  training  data  size} 
  We analyze the
  overall  improvement  trend  of  SFC under various few-shot settings.

  Our main
  focus  is  to  study  chatbot  building  in common situation during real-world
  application  where  only  a few sample sentences are available for each class.
  The  experimental  results  for  the  3  diverse  intention detection datasets
  (ClINC150,  BANKING77,  HWU64)  under  few-shot  settings, 5 to 20 samples per
  class,  indicate that our proposed joint SFC can achieve 2.47 percents average
  improvement   over  one  of  the  most  powerful  baseline  models,  which  is
  RoBERTa-large  classification  model.  Moreover,  the  improvement  in F-score
  becomes  more  and  more prominent with the decrease of data sample number per
  class.

  Especially,  in the most extreme few-shot setting with only 5 training samples
  per  class  our joint SFC achieves 4.97 percent points improvements on average
  over  RoBERTa-large classification model . This observation indicates that SFC
  is  especially  well-performed  when being applied during the initial stage of
  building  task-specific  chatbot  system,  where the amount of data sample for
  each class label is extremely scarce.

  Furthermore,  our  proposed  joint  SFC can also achieve 2.04 percents average
  improvement  over  RoBERTa-large classification model on all the data settings
  including  the  ones  with  large  scale,  30  to  50  samples per class. This
  indicates that joint SFC can still steadily outperforms pretrained model based
  baseline even if the data size grows bigger. However, the improvement is still
  more prominent when the data is scarce, which makes SFC an excellent model for
  low-resource chatbot scenario.

  \subsubsection*{How is joint SFC related to the label  embedding based  LEAM?}
  LEAM performs consistently better than TextCNN by a large margin in all
  datasets. TextCNN is practically efficiently in task-specific chat
  applications, and LEAM shows its power in modeling labeling information.

  LEAM  does not perform as well as SFCs, since the first intuitive reason is it
  does  not use pretrained model as encoding module; and another critical reason
  is  in  task-specific  chat applications, it is common to have many intentions
  that are close to some others with a minor difference, and it is hard and even
  impossible  to  name  each  intention  with  a short clear name. One candidate
  solution  is  setting  a most standard sample as the label of that class. When
  using  non-pretrained  models  base  LEAM,  this is applicable. Yet when using
  pretrained  model  based, as the all labels can be hundreds of thousands, then
  this  can not be accommodated in a poplar 32 G Tesla V100 GPU. Actually, joint
  SFC  can  be  kind  of  understood  as  a  generalization form of LEAM. In the
  scenario when there is no clear class label, the relationship between a sample
  and  a class lable is implicatily encoded as that between a sample and another
  sample  from  the  same  class, and this turns into a sentence pair similarity
  model.

  \subsubsection*{Settings  of  hyperparameters} 
  Finally, we analyze the effect brought
  by  different  settings  of  hyperparameters,  see  Table \ref{tbe:table3}.

  We  conduct  experiments on 5 different settings of hyperparameters, which are
  the  number of candidate class number $K$ and the number of sentence pairs $P$
  for  each  class,  on  ITG  and Amazon-670K datasets to find the most suitable
  setting  of  these  2 hyperparameters. We control the total number of sentence
  pairs  sampled for each user query into about the same level of amount(50-60).
  Through  analyzing the experimenal results shown in Table \ref{tbe:table3}, we
  find  that the best setting of hyperparameters might be different for datasets
  of  different  data  size.  The  class label number of ITG and Amazon-670k are
  similar  in  our  experimental setting, which is around 250. However the total
  amount of data points of ITG is approximately 2 times larger than Amazon-670k.
  Therefore,  the  setting of $K=5$ and $P=10$ is good enough for ITG since most
  of  the  classes  contains  over 10 data points. However, for Amazon-670k, the
  amount  of  data  sample  for  each  class  is  scarce, thus, we may need more
  candidate  classes to sample enough effective sentence pairs for sentence-pair
  model  layer  of  SFC.  In this way, the setting of $K=10$ and $P=5$ becomes a
  good choice. However, in spite of the effect on evaluation performance brought
  by different settings of hyperparameters, joint SFC still steadily outperforms
  two-stage  SFC  by  0.41  percents  in  overall  average  F-score  on  ITG and
  Amazon-670k datasets.
